### 2023-11-14 manuscript/manuscript.tex
Continuation of discussion

\documentclass[12pt]{article}

\usepackage{amsmath}
\usepackage[margin = 1in]{geometry}
\usepackage{graphicx}
\usepackage{booktabs}
\usepackage{natbib}
\usepackage[colorlinks=true, citecolor=blue]{hyperref}

\title{Predicting Diabetes Based on Various Factors Using logistic Regression}
\author{Ashley Merritt\\
    Department of Statistics\\
    University of Connecticut
    }

\begin{document}
\maketitle

\begin{abstract}

\end{abstract}

The rest of the paper is organized as follows.
The data will be presented in Section~\ref{sec:data}.
The methods are described in Section~\ref{sec:meth}.
The results are reported in Section~\ref{sec:resu}.
A discussion concludes in Section~\ref{sec:disc}.

\section{Introduction}
\label{sec:intro}
    According to the American Diabetes Association, about 11.3 percent of Americans, or about 37.3 million people, have diabetes. Of those 37.3 million people, about 28.7 million were actually diagnosed with diabetes, while the remaining 8.6 million were left undiagnosed (\citet{CDC2022Diabetes}). Those left undiagnosed are at risk for even more serious illness if left untreated. Diabetes is a disease where your body does not create enough insulin, or use it properly, in order to get glucose into your cells and use it for energy (\cite{NIH2023Whatis}). Therefore, it is very important that your body has a normal glucose level. In this paper, I will be working to determine the pre-existing factors that can be used in order to predict if a person may have diabetes. Establishing the predictors is very important to me as many of my family members have a long history of suffering from this disease. It is important to better predict this disease to save people from unnecessary suffering.

    Previously, the prediction of diabetes has been studied using machine learning models. Specifically in Sisodia's paper, it focused on the "Prediction of Diabetes using Classification Algorithms", they worked through the topic by producing results using support vector machine, Naive Bayes classifier, and decision tree classifier (\cite{Sisodia2018Prediction}). In this paper, I will be continuing the investigation with a different data set using logistic regression and various other statistical tests..

    In this paper the specific research question that I will be focusing on is: how can we use machine learning models in order to identify individuals that may have already or are at risk at developing diabetes? This will be researched using a data set that contains various metrics on one's health that will be described in the next section of the paper.

\section{Data}
\label{sec:data}
    In order to study the proposed research question on the topic of diabetes, I searched through many databases in order to find a set that had all of the information I was looking for. After exploration, I found a data set entitled "Healthcare Diabetes Data set" on the website Kaggle. The data set was originally sourced from the National Institute of Diabetes and Digestive and Kidney Diseases. The data encompasses eight different predictors including pregnancies, glucose, blood pressure, skin thickness, insulin, BMI, diabetes pedigree function, and age.

    The predictors listed above will be represented in my research as the following variables. 'Pregnancies' will provide the number of times an entry has been pregnant. 'Glucose' will provide the plasma glucose concentration over two hours using the results of an oral glucose tolerance test on a entry. 'BloodPressure' gives the diastolic blood pressure in mm Hg of an entry. 'SkinThickness' will provide the triceps skin fold thickness in mm. 'Insulin' will include a test on two hour serum insulin in mu U/ml. 'BMI' will provide a calculation of weight (kg) divided by height ($m^2$). 'DiabetesPedigreeFunction' will provide a genetic score of diabetes. 'Age' will simply be the entry's age at the time of the collection. Following this paragraph, I will provide the link to the dataset that I will be using as well as showing some descriptive statistics in Table~\ref{tab:ds} including the mean, standard deviation, and median.

Dataset link: \href{https://www.kaggle.com/datasets/nanditapore/healthcare-diabetes}{Healthcare Diabetes}

\begin{table}[ht]
    \caption{Descriptive Statistics of Healthcare Diabetes Dataset}
  \label{tab:ds}
\centering
\begin{tabular}{lrrr}
      \hline
    Variable & Mean & SD & Median \\ 
      \hline
      Pregnancies & 3.74 & 3.32 & 3.00 \\ 
      Glucose & 121.10 & 32.04 & 117.00 \\ 
      BloodPressure & 69.13 & 19.23 & 72.00 \\ 
      SkinThickness & 20.82 & 16.06 & 23.00 \\ 
      Insulin & 80.13 & 112.30 & 37.00 \\ 
      BMI & 32.14 & 8.08 & 32.20 \\ 
      DiabetesPedigreeFunction & 0.47 & 0.33 & 0.38 \\ 
      Age & 33.13 & 11.78 & 29.00 \\ 
       \hline
    \end{tabular}
    \end{table}

\section{Methods}
\label{sec:meth}
    Various methods will be used to complete the statistical analysis necessary for this project. I will begin by exploring basic descriptive statistics to gain an understanding of my data set. I will then progress to creating visualizations including histograms and box plots to understand relationships and patterns within my data. In order to draw insight on my data I will perform logistic regression analysis to determine which predictors are significant in predicting if someone has diabetes. In order to perform the logistic regression, I will have to check the necessary assumptions associated with the method. Within this analysis I will perform necessary transformations in order to create the best model that I can. From my results I will draw necessary conclusions about my data.

    The first assumption that will be checked in that the dependent variable is binary, this true as our outcome variable is in the form of 1 (having diabetes) or 2 (not having diabetes). The next assumption that will be checked is if the observations are independent of each other, this is also true for our data as each predictor is calculated or taken independent of one another. The third assumption that will be checked is that there is no multicollinearity among the independent variables. In order to check this assumption we will calculate the variance inflation factor, VIF, in order to measure the strength of the correlation between predictor variables in our model. One example of this is seen in Table~\ref{tab:vif} where all of the values are around 1, indicating that there is moderate correlation between pregnancies and the other predictors but this value is not in a concerning range.

\begin{table}[ht]
\caption{VIF Values} 
\label{tab:vif}
\centering
\begin{tabular}{rr}
  \hline
 & vif(lm(P \~{} G + BP + ST + I + BMI + DPF + A)) \\ 
  \hline
Glucose & 1.29 \\ 
  BloodPressure & 1.18 \\ 
  SkinThickness & 1.50 \\ 
  Insulin & 1.44 \\ 
  BMI & 1.30 \\ 
  DiabetesPedigreeFunction & 1.06 \\ 
  Age & 1.18 \\ 
   \hline
\end{tabular}
\end{table}

    After exploring the descriptive statistics, I will begin to perform a logistic regression. Logistic regression, is used to describe the data and the relationship between one dependent variable against many independent variables. In this particular case, we will be looking at the binary response of diabetes with the various independent variables including but not limited to; pregnancies and glucose levels. To begin, I will split the data into a training and a test set. This will be done by setting a random seed of 30 and using 80 percent of the data set in the training set and the remaining 20 percent as the testing set. We will next begin to build the logistic regression model using the training data.

    In Table~\ref{tab:sl} you will see the output of each predictor on predicting the outcome of diabetes. This table will provide us with coefficients, the standard errors, the z-scores, and finally the p-value for each predictor in the model. Based on the output, we can determine that only skin thickness is not statistically significant in predicting if a patient may have diabetes. Therefore, all of the other predictors are statistically significant and should be monitored in predicting diabetes.

\begin{table}[ht]
    \caption{Summary of Logistic Regression Model}
    \label{tab:sl}
\centering
\begin{tabular}{lrrrr}
  \hline
Predictor & Estimate & Std. Error & z value & Pr($>$$|$z$|$) \\ 
  \hline
(Intercept) & -8.02 & 0.41 & -19.73 & 0.00 \\ 
  Pregnancies & 0.11 & 0.02 & 6.04 & 0.00 \\ 
  Glucose & 0.03 & 0.00 & 16.34 & 0.00 \\ 
  BloodPressure & -0.01 & 0.00 & -3.69 & 0.00 \\ 
  SkinThickness & 0.00 & 0.00 & 0.26 & 0.79 \\ 
  Insulin & -0.00 & 0.00 & -2.74 & 0.01 \\ 
  BMI & 0.08 & 0.01 & 9.07 & 0.00 \\ 
  DiabetesPedigreeFunction & 0.76 & 0.17 & 4.38 & 0.00 \\ 
  Age & 0.02 & 0.01 & 3.03 & 0.00 \\ 
   \hline
\end{tabular}
\end{table}
\section{Results}
\label{sec:resu}

After completing our methods section, we can now move on to talk about the results of our model. Based on our logistic regression model, we now know that all of the predictors in the data set except skin thickness are important in order to predict whether a patient may have diabetes or not. With this information medical professionals will be able to ensure proper treatment to patients as they can use these predictors in order to tell if a patient may have diabetes even sooner than typical means. The model allows for medical professionals to use the model in order to further their research into diabetes and the various precursors.

\section{Discussion}
\label{sec:disc}

Currently, this study is limited as we are set within the bounds of this data set. In future studies we can include more predictors in order to see if there are any other factors that could be used to predict diabetes. As well as having an even larger data set in order to ensure our accuracy. Other data sets may suggest that certain predictors are not helpful in predicting diabetes that our model did not pick up on due to the metrics of our data.

As this is a paper dealing with sensitive medical data, we have many limitations in order to be ethical as well as respecting patient privacy. Based on the data set collected we should be able to ensure patient privacy pretty well as names or any very unique identifiers are not listed with each entry. We simply have a ID and age for each entry so we do not have to worry as much for the sake of privacy. In terms of our direct models we may have to disregard some assumptions as they may not fit with real world data. If something unexpected happens we may have to reconsider the data set used and consider finding a new one. This will be done by a case by case scenario and will be investigated thoroughly before making any decisions.

\bibliography{refs}
\bibliographystyle{mcap}

\end{document}